Weakly informative observations: not the winner

### 2020-04-13 inst/notthewinner.Rmd
fixes #51

Here we consider a situation in which we make an observation that is
only very very weakly informative.  We consider the observation
${\mathcal O}$ that a particular player *does not* win: the order
statistic is unobserved and we know nothing about it except for the
fact that (wlog) player 1 (aka "me") is not first placed.

As in `loser.Rmd`, we have three hypotheses:

* $H_0\colon\sum p_i=1$  (that is, no restriction)
* $H_1\colon\exists i\neq 1 \mbox{ with } p_1\geq p_i>0$
* $H_2\colon p_1=p_2=\ldots=p_n=\frac{1}{n}$

(see how $H_1$ implies that someone is worse than me).  Observe that
$H_1$ and $H_2$ correspond to a restriction on allowable parameter
space.

To understand the informativeness of ${\mathcal O}$, we will consider
it in conjunction with some highly informative observations ${\mathcal
I}$ which are order statistics:

@@

We will firstly consider these observations in isolation, then later
combine them with ${\mathcal O}$ which we will take to be
(conditionally) independent.  The first step is to define a suitable
likelihood function for ${\mathcal I}$:q

@@

Above, `mH0` corresponds to $H_0$.  Note the low but nonzero estimated
strength of `p1`.

Observe in passing that we may reject $H_z\colon p_1=0$
unconditionally: we can deduce using elementary arguments that the
likelihood of $H_z$ is zero: if $p_1=0$, then $p_4=0$ from the first
line, and then $p_1=0$ from line 2; then $p_3=0$ from line 3, $p_5=0$
from line 4 and finally $p_2=0$ from line 1.  Thus we can say for sure
that $H_z$ is incorrect and we may safely ignore it henceforth.

## Hypothesis 1

Hypothesis 1 is rather tricky so we will use the ersatz $H_{1'}\colon
p_1\geqslant p_4$ (we choose $p_4$ on the grounds of its low estimated
value).  Then

```{r}
n <- 5 # number of players
small <- 1e-3  # numerical necessity
(mH1 <- maxp(H,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0))
loglik(indep(mH1),H)
```

See how the constraint of $H_1$ is active, for at the evaluate
$p_1=p_4$ (the numerical value of `small` is immaterial).

## Hypothesis 2

```{r}
mH2 <- equalp(H)
loglik(indep(mH2),H)
```

## Support function

We can present a support function for $H_0,H_1,H_2$ for ${\mathcal
I}$:

```{r,suppcalculator}
Supp <- c(H0=loglik(indep(mH0),H), H1=loglik(indep(mH1),H), H2=loglik(indep(mH2),H)
)
Supp-max(Supp)
```

@@

```{r,mls0}
mSL0 <- maxp(SL)
mSL1 <- maxp(SL,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0)
mSL2 <- equalp(SL)
```

@@

which is identical to the previous support function.

@@

```{r,mlsa}
mSLa0 <- maxp(SLa)
mSLa1 <- maxp(SLa,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0)
mSLa2 <- equalp(SLa)
```

>>>>

Here we consider a situation in which we make an observation that is
only very very weakly informative.  We consider the observation
${\mathcal O}$ that a particular player *does not* win: the order
statistic is unobserved and we know nothing about it except for the
fact that (wlog) player 1 ($p_1$, aka "me") is not first placed.  We
consider whether there is evidence to support two different assertions:

* $H_1\colon p_1=p_2=\ldots=p_n=\frac{1}{n}$
* $H_2\colon\exists i\neq 1 \mbox{ with } p_1\geq p_i>0$

Our alternative would be that $(p_1,\ldots,p_n) is unconstrained
except for the unit sum constraint:

* $H_0\colon\sum p_i=1$  (that is, no restriction)

(see how $H_2$ implies that someone is worse than me).  Observe that
$H_1$ and $H_2$ correspond to a restriction on allowable parameter
space.  To understand the informativeness of ${\mathcal O}$, we will
consider it in conjunction with some highly informative observations
${\mathcal I}$ which are order statistics:

@@

We will firstly consider these observations in isolation, then later
combine them with ${\mathcal O}$ which we will take to be
(conditionally) independent.  The first step is estimate the players'
strengths:

@@

Above, `mH0` corresponds to an unconstrained optimization.  Note the
low but nonzero estimated strength of `p_1`.  Observe in passing that
we may reject $H_z\colon p_1=0$ unconditionally: we can deduce using
elementary arguments that the likelihood of $H_z$ is zero: if $p_1=0$,
then $p_4=0$ from the first line [because $p_4$ beat $p_1$ on that
occasion], and then $p_1=0$ from line 2; then $p_3=0$ from line 3,
$p_5=0$ from line 4 and finally $p_2=0$ from line 1.  Thus we can say
for sure that $H_z$ is incorrect and we may safely ignore it
henceforth.

## Hypothesis 1: equality

```{r}
equalp.test(H)
```

Thus there is no evidence to suggest that the players have different
strengths: we are unable to reject $H_1$, the hypothesis of equality.
Hypothesis $H_2$ is rather tricky so we will use the ersatz
alternative that $p_1\geqslant p_4$ (we choose $p_4$ on the grounds of
its low estimated value).  Then

```{r}
n <- 5 # number of players
small <- 1e-3  # numerical necessity
(mH1 <- maxp(H,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0))
loglik(indep(mH1),H)
loglik(indep(mH0),H)-loglik(indep(mH1),H)   # support
```

There is no reason to reject our null that $p_1\geqslant p_4$, for the
support falls below the two units of support criterion.  Further, see
how the constraint is active, for at the evaluate $p_1=p_4$ (the
numerical value of `small` is immaterial).

@@

```{r,mls0,cache=TRUE}
mSL0 <- maxp(SL)
mSL1 <- maxp(SL,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0)
mSL2 <- equalp(SL)
```

@@

which is identical to the previous support function for $H_1$ and
$H_2$ (apart from a traditional sign) taken from `equalp.test()` and
`mSL2` above.

@@

```{r,mlsa,cache=TRUE}
mSLa0 <- maxp(SLa)
mSLa1 <- maxp(SLa,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0)
mSLa2 <- equalp(SLa)
```
